### 2015-01-15 docs/rebuttal.txt
work on the first reviewer

- General Remarks:

We want to thank the reviewers for their meaningful comments.

- Reviewer 1:

# dimitro: I do not know exactly what to say about the following comment:
# "The sensitivity and specificity metrics are questionable (need
# justification and relation to standard techniques), and anyway not
# used much in Table I."
# What does he mean by justification? We provide the corresponding citations
# for the metrics. Finally, we can say that we can say that ce can calculate
# them and add them to the Table.

# dimitro: I do not know what to say about this...
# "The paper is tutorial-like and broad, which is not so bad by itself,
# but it is desirable of a SoK paper to contain more insight than a
# typical survey."

It is true that there are to many N/As in both Tables and especially in
Table II. However, we believe that issues concerning code availability and
accuracy testing are instantly identified by a casual reader.

We are planning to change the paper's title into: "Defending Against Code
Injection Attacks that target Web Applications: From Research Approaches
to Practical Tools".

# dimitro: or ""Defending Against Code Injection Attacks on the Web:
# From Research Approaches to Practical Tools"."
# If you have any other ideas let me know.

As we mention in the paper, we analyze and categorize mechanisms that have
been described in research papers. Still, we can include ADsafe,
Facebook JavaScript, and Secure EcmaScript in the paper in the way
we include Google Caja. In addition, we agree that we should include
FlowFox and JSand (both have >20 citations which is our threshold),
and the paper presenting IFC in WebKit's JavaScript Bytecode (which is
a recent work presented in a top venue). We could also include AdSentry, Contego
and JSFlow too but currently, they do not reach the citation threshold.

# dimitro: is POST (Conference on Principles of Security and Trust)
# a top venue?

We agree that some of the presented approaches utilize either static or
dynamic analysis. As we explain in the paper though, we dismiss the
static and dynamic approaches that identify vulnerabilities that
may lead to the attacks that we refer to and cover protection
mechanisms that stop such attacks when they are launched.

# dimitro: I do not fully understand the following:
# "The security discussion (Section IV) is somewhat minimalistic.
# Further, more insight would be helpful on the security of the systems
# and the attack surfaces covered beyond what is reported in Table I
# (which talks about the classical SQL injection, XSS, CSRF and known
# DOM-based, JSON injection, and XPath injection attacks)."

We can update our model to show where the language-level and web
framework-level protection mechanisms detect an attack. In particular
we can show that this is done at the application level within the server.

The comments concerning Firing Range, XCS attacks, scriptless attacks
and attacks that involve plug-ins are totally valid and we are
planning to address them in the next version of the paper.

- Reviewer 2:

- Reviewer 3:

- Reviewer 4: