### 2015-01-15 docs/reviewers_response.txt
add PL reviewers' response

Reviewer 1
==========

> -The sensitivity and specificity metrics are questionable (need
> justification and relation to standard techniques), and anyway not
> used much in Table I.

That's exactly what we try to highlight in the paper. Sensitivity and
specificity should be more widely applicable, but they are not,
because not enough data is given in the original publications. In
general, as we remark in VII.A, "In many cases researchers tend not to
provide the false positive and negative results that their tools may
produce."

Concerning the usefullness of sensitivity and specicifity as metrics
per se, their use in computer security has been advocated by other
researchers as well, for example ***.

Reviewer 2
==========

> The paper essentially provides only information that is directly
> available from other publications. This applies especially to the
> categories accuracy and performance overhead, where several mechanisms
> do not provide data. The authors of the paper did not perform any
> original tests.

Please note that even if we wanted to perform original tests on the
systems we studied, or even if we wanted to validate the existing
tests, we would not be able to do so in the vast majority of cases, as
indicated in Table II; this is in fact one of our main findings, as we
explain in VII.B.

Reviewer 3
==========

> From my understanding, sensitivity and specificity could also be
> expressed in terms of conditional probability: sensitivity is the
> conditional probability that the detection mechanism flags an attack,
> given that an attack is taking place, and specificity is the
> conditional probability that the mechanism does not report an attack,
> given that no attack takes place.

Correct. We can of course add that to the power; we just thought that
it might not be necessary to explain it.

> This would help in discerning and relating these measures from and to
> the positive and negative predictive values, and also explains
> formulas (5) and (6) on page 13.

Correct, as above we can easily add the explanations.

> Moreover, the discussion seems to be confusing experiment measurements
> and abstract probabilities.
> On page 2 [actually it's on p. 5], it is stated that "sensitivity
> and specificity can be calculated on test data alone".
> On page 13, it is said that "PPV and NPV ... relate to the
> effectiveness in an actual production environment".

This is to highlight that we cannot carry out a set of measurements,
calculate sensitity and specificity, and then using the {True | False
} {Positives | Negatives} calculate the PPV and the NPV from equations
3 and 4 directly. One has to go instead through equations 5 and 6.

> Obviously, any measurement of true and false positives and negatives
> on test data can be skewed, if the test data is skewed.
> Hence, if the sensitivity and specificity are measured using tests
> that do not relate to "the production environment", the calculated
> values are meaningless.

Definitely, we can highlight that.

> Yet this holds true for all four measures; it is simply related to the
> way they are estimated.
> The advantage of PPV and NPV is that they answer the question of
> "given a detection result, how much can you trust it?", which is more
> interesting in the "production environment".
>
> This part of the paper should be revised to give a clearer picture.

Taking into account all your related comments, we will do that, taking
care not to go through the paper length limits. The gist is that,
indeed, PPV and NPV answer the question you state; it is not always
possible to get data for them (since we may not know the prevalence),
but it may worth to try and get it, or this could be a new research
direction. Of course, if we do have such data, it is a way to
operationalize the protection a system provides.

The validity of test data is in fact important independently of
sensitivity and specificity; skewed test data will give a wrong
picture anyway, even if nobody bothers to calculate sensitivity and
specificity. We do not know in the papers we studies whether test data
were skewed. We do find, however, that: more tests could have been
carried out (cf. Table I and VII.A ) and the means to conduct tests should be
made available to third parties (cf. Table II and VII.B).

Reviewer 4
==========

> Similarly, the demand on taking the prevalence into consideration when
> evaluating false positives and false negatives may be an undue burden
> on security researchers, who would have a hard time collecting these
> probabilities. On the other hand, this might call for a common attack
> dataset.

That is why we state, in p. 13, that "With this in mind it may be
unfair to ask of researchers to provide PPV and NPV values for
their mechanisms---in fact, we see that nobody does." In the following
paragraph (last paragraph of first column in p. 13) we explore what
this would require. We can make it more clear.