### 2015-01-16 docs/rebuttal_merged.txt
664 words

We thank the reviewers for their insightful comments.

Overall, we didn't consider mechanisms that
haven't been presented in research papers. We mention Caja
due to its importance.

Although some of the presented works use
static/dynamic analysis, we don't
consider the broader area of static/dynamic analysis for
vulnerability detection (another paper covers that area [14]),
but focus only on protections against attacks
that exploit such vulnerabilities.

We will update our paper
to address the issues indicated by the
the reviewers, i.e. the title,
the point where language protection mechanisms detect an attack,
the "ease of use" dimension, the SQLi example,
the fact that we fail to capture PHP injection in our model,
the 1/1.1 notation in our model, and
the code availability discussion.

Rev1:

Sensitivity and specificity should be more widely applicable,
but not enough data is given in the publications (see VII.A).
The usefulness of such metrics in computer security
has been advocated by other researchers too, i.e.
Pfleeger and Pfleeger (Analyzing Computer Security:
A Threat/Vulnerability/Countermeasure Approach,
Prentice Hall, 2011); we’ll add the citations.

There are too many N/As, especially Table II, but this actually
highlights the issues regarding code availability and accuracy
testing.

We will include FlowFox and JSand (both have >20 citations),
the paper presenting IFC in WebKit's
JavaScript Bytecode (a recent paper)
and discuss ADsafe etc. in the same way as Caja.
AdSentry, Contego, and JSFlow 
don't reach our citation threshold.

We agree that using "security" as a dimension
is misleading---we use it in the sense of "bypassability", which is 
orthogonal to "accuracy". We will replace it with a more specific term.

We agree about the minimalism of the security discussion, and
will try to enrich this part. For instance, approaches that provide
systematic arguments on why their designs are secure, aren't as
"bypassable" as heuristic/threshold-based approaches.

Rev2:

The lack of code availability for most systems makes performing
original tests challenging, which is one of our findings, as we
explain in VII.B.

We attempt to equally cover every mechanism and then
highlight the reasons why a mechanism may be widely
used or not. In our conclusion section, we particularly mention
that CSP is one of the few that 
has been published and is widely used.
It is true that if programmers use
secure coding practices like prepared statements, then
SQLi protections aren't required.
However, best programming practices aren't always followed,
and this motivates proposals like SQLrand,
which might be easier to retrofit in existing code bases.

Our model covers attacks that utilize DSL languages in
general, and in our categorization there are more than
one countermeasures that protect from XPathi.

Rev3:

Of course sensitivity and specificity could be expressed in terms of a
conditional probability---we will include this in the paper.

Our discussion in pages 2 and 13 highlights that we
cannot perform a set of measurements, calculate sensitivity and
specificity, and then using {True|False} {Positives|Negatives},
calculate PPV and NPV from equations 3 and 4 directly. One has to go
instead through equations 5 and 6.

We will expand on the advantages of PPV and NPV. PPV and NPV answer
the question about trusting a detection result; however, it isn't
always possible to get data for them as we indicate in p.13.

The comment on the skewed data is valid. The validity of test data is
important independently of sensitivity and specificity. We don't know
in the papers we studied if test data were skewed. We do find,
though, that: more tests could have been performed (Table I and
VII.A) and the means to conduct tests should be made available to
third parties (Table II and VII.B).

Rev4:

We don't intend to criticize the lack of testing
in etiological mechanisms. We attempt to 
explore why such mechanisms aren't widely used, and suggest
that this is one reason.

Regarding the comment about demanding prevalence, it may be unfair to
ask of researchers to provide PPV and NPV values for their mechanisms.
In p.13 we explore what this would require---we will elaborate
more.