### 2015-01-16 docs/rebuttal_merged.txt
reduce further

We thank the reviewers for their insightful comments.

Some general points:

* We didn't consider mechanisms that haven't been presented in
research papers. We mention Caja due to its importance.

* Although some of the presented works use static/dynamic analysis, we
 don't consider that area (another paper covers it [14]), but
 focus only on protections against attacks that exploit such
 vulnerabilities.

* We will update our paper to address the issues indicated by the the
  reviewers, i.e. the title, the point where language protection
  mechanisms detect an attack, the "ease of use" dimension, the SQLi
  example, the fact that we fail to capture PHP injection in our
  model, the 1/1.1 notation in our model, and the code availability
  discussion.

Rev1:

Sensitivity and specificity should be more widely applicable, but not
enough data is given in the publications (see VII.A). The usefulness
of such metrics in computer security has been advocated by others,
i.e. Pfleeger and Pfleeger (Analyzing Computer Security: A
Threat/Vulnerability/Countermeasure Approach, Prentice Hall, 2011);
we’ll add the citation.

There are too many N/As in Table II, but this actually highlights the
issues regarding code availability and accuracy testing.

We will include FlowFox and JSand, the paper presenting IFC in
WebKit's JavaScript Bytecode and discuss ADsafe etc. and Caja.
AdSentry, Contego, and JSFlow don't reach our citation threshold.

Using "security" as a dimension is misleading---we mean ease of
circunvention, which is orthogonal to "accuracy". We
will replace it with a more specific term.

We'll enrich the security discussion. For instance, approaches that
provide systematic arguments on why their designs are secure, cannot
be circumvented like heuristic/threshold-based approaches.

Rev2:

The lack of code availability for most systems makes performing
original tests challenging, which is one of our findings (VII.B).

We attempt to cover mechanisms and then highlight reasons why a
mechanism may not be widely used. In the conclusions we mention that
CSP is one of the few that has been published and is widely used. It
is true that if programmers use secure coding practices like prepared
statements, then SQLi protections aren't required. However, that may
not happen, and this motivates proposals like SQLrand, which might be
easier to retrofit in existing code bases.

Our model covers attacks that utilize DSL languages in general, and in
our categorization there are more than one countermeasures that
protect from XPathi.

Rev3:

Sensitivity and specificity could be expressed in terms of a
conditional probability---we will include this in the paper.

Pages 2 and 13 highlight that we cannot perform a set of measurements,
calculate sensitivity and specificity, and then using {True|False}
{Positives|Negatives}, calculate PPV and NPV from equations 3 and 4
directly. One has to go instead through equations 5 and 6.

We will expand on the advantages of PPV and NPV. PPV and NPV answer
the question about trusting a detection result; however, it isn't
always possible to get data for them as we indicate in p.13.

The comment on the skewed data is valid. We don't know in the papers
we studied if test data were skewed. We do find, though, that: more
tests could have been performed (Table I and VII.A) and the means to
conduct tests should be made available to third parties (Table II and
VII.B).

Rev4:

We don't intend to criticize the lack of testing in etiological
mechanisms. We attempt to explore why such mechanisms aren't widely
used, and suggest that this is one reason.

Regarding the comment about demanding prevalence, it may be unfair to
ask of researchers to provide PPV and NPV values for their mechanisms.
In p.13 we explore what this would require---we will elaborate.